---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nonprobsvy an R package for modern inference methods based on nonprobability samples.

<!-- badges: start -->
  [![R-CMD-check](https://github.com/ncn-foreigners/nonprobsvy/workflows/R-CMD-check/badge.svg)](https://github.com/ncn-foreigners/nonprobsvy/actions)
<!-- badges: end -->

## Basic information

This package implements mass imputation, inverse probability weighting and doubly robust estimators based on the following papers:

+ Yang, S., Kim, J. K., & Song, R. (2020). Doubly robust inference when combining probability and non-probability samples with high dimensional data. Journal of the Royal Statistical Society. Series B, Statistical Methodology, 82(2), 445.
+ Kim, J. K., Park, S., Chen, Y., & Wu, C. (2021). Combining non-probability and probability survey samples through mass imputation. Journal of the Royal Statistical Society Series A: Statistics in Society, 184(3), 941-963.
+ Chen, Y., Li, P., & Wu, C. (2020). Doubly robust inference with nonprobability survey samples. Journal of the American Statistical Association, 115(532), 2011-2021.

## Installation

You can install the development version of `nonprobsvy` package from [Github](https://github.com/ncn-foreigners/nonprobsvy) with:

```{r, eval=FALSE}
remotes::install_github("ncn-foreigners/nonprobsvy")
```

## Example

Simulate example data from the following paper: Kim, Jae Kwang, and Zhonglei Wang. "Sampling techniques for big data analysis." International Statistical Review 87 (2019): S177-S191 [section 5.2]

```{r example}
library(survey)
library(nonprobsvy)

set.seed(1234567890)
N <- 1e6 ## 1000000
n <- 1000
x1 <- rnorm(n = N, mean = 1, sd = 1)
x2 <- rexp(n = N, rate = 1)
epsilon <- rnorm(n = N) # rnorm(N)
y1 <- 1 + x1 + x2 + epsilon
y2 <- 0.5*(x1 - 0.5)^2 + x2 + epsilon
p1 <- exp(x2)/(1+exp(x2))
p2 <- exp(-0.5+0.5*(x2-2)^2)/(1+exp(-0.5+0.5*(x2-2)^2))
flag_bd1 <- rbinom(n = N, size = 1, prob = p1)
flag_srs <- as.numeric(1:N %in% sample(1:N, size = n))
base_w_srs <- N/n
population <- data.frame(x1,x2,y1,y2,p1,p2,base_w_srs, flag_bd1, flag_srs)
base_w_bd <- N/sum(population$flag_bd1)
```

Declare `svydesign` object with `survey` package

```{r}
sample_prob <- svydesign(ids= ~1, weights = ~ base_w_srs, 
                         data = subset(population, flag_srs == 1))
```

Estimate population mean of `y1` based on doubly robust estimator using IPW with calibration constraints.

```{r}
result_dr <- nonprob(
  selection = ~ x2,
  outcome = y1 ~ x1 + x2,
  data = subset(population, flag_bd1 == 1),
  svydesign = sample_prob
)
```

Results

```{r}
summary(result_dr)
```

Mass imputation estimator 

```{r}
result_mi <- nonprob(
  outcome = y1 ~ x1 + x2,
  data = subset(population, flag_bd1 == 1),
  svydesign = sample_prob
)
```

Results

```{r}
summary(result_mi)
```

Inverse probability weighting estimator

```{r}
result_ipw <- nonprob(
  selection = ~ x2,
  target = ~y1,
  data = subset(population, flag_bd1 == 1),
  svydesign = sample_prob)
```

Results

```{r}
summary(result_ipw)
```

## Funding

Work on this package is supported by the the National Science Center, OPUS 22 grant no. 2020/39/B/HS4/00941.
